---
title: "Microbiome Project"
author: "Tony Chen"
date: "11/9/2021"
output:
  word_document: default
  html_document: default
---



# Input the microbiome data: 38 subjects, 82 microbes. These data have undergone log-ratio transfomation and so can be deemed as normal data. As for the Outcome data also is input in a separate file.

```{r setup,echo=F, include=FALSE}
require(data.table)
require(dplyr)
require(magrittr)
require(tidyr)
library(CVglasso)
require(igraph)
require(ggplot2)
require(intergraph)
require(GGally)
library(sna)
require(glasso)
require(glmnet)
require(corrplot)

#taxa_all = read.csv("")
taxa_raw = read.csv("C:/Users/Jie Zhou/Documents/paper02052022/longitudinalOutcome/Cystic-Fibrosis-and-Gut-Microbiome/data/ddata.csv")
pulm = read.delim("C:/Users/Jie Zhou/Documents/paper02052022/longitudinalOutcome/Cystic-Fibrosis-and-Gut-Microbiome/data/analyst_pulmonary_exacerbation_202110061618.csv")
variable.names(pulm)

# averaging across time points
taxa_avg = taxa_raw %>% 
  group_by(subject) %>% 
  summarize_all(list(mean))

taxa = taxa_avg[,4:ncol(taxa_avg)]
```


# Selecting the microbial interaction network using glasso. Here the tuning parameter is recommended based on previous study. Given the selected network, leadeigen algorithm is employed to find the communities in the network. 

```{r,echo=F}
set.seed(2)
# lambda_glasso = seq(0.05, 6, 0.01)
# cv_glasso_taxa = CVglasso(taxa, K = 5, lam = lambda_glasso)
#   lam_pick = cv_glasso_taxa$Tuning[2]
lam_pick=0.75
### asymptotic distribution of hyper parameter
# lam_pick = rep(NA, 1)
# for (i in 1 : 1) {
#   cv_glasso_taxa = CVglasso(taxa, K = 5, lam = lambda_glasso, trace = "none")
#   lam_pick[i] = cv_glasso_taxa$Tuning[2]
#   print(i)
# }
# hist(lam_pick)
# lam_pick = as.data.frame(lam_pick)
# my_lambdas = read.csv("lambdas.csv")
###

#glasso_taxa = glasso(cov(taxa), rho = cv_glasso_taxa$Tuning[2])
glasso_taxa = glasso(cov(taxa), rho = lam_pick)
pcov_taxa = glasso_taxa$wi
#pcov_taxa = cov2cor(pcor_taxa)
taxa_nwk = as.matrix(ifelse(pcov_taxa != 0, 1, 0))
#corrplot(taxa_nwk, type = "upper", diag = T, tl.col = "white")

# graph theory

taxa_graph = graph.adjacency(taxa_nwk, mode = "undirected")
taxa_louvain = cluster_louvain(taxa_graph)
taxa_leadeigen = cluster_leading_eigen(taxa_graph) # This one produces fewer modules (clusters)


#plot(simplify(taxa_graph), vertex.label=NA, vertex.size=5, vertex.color = taxa_louvain$membership)
#ggnet2(taxa_graph, color = taxa_louvain$membership, size = "degree")
#ggnet2(taxa_graph, color = taxa_leadeigen$membership, size = "degree")
```

# For each of the communities, compute its principal components.





# For each of the communities, compute its principal components.
```{r,echo=F}
colnames(taxa[which(taxa_leadeigen$membership == 3)])

taxa_le_cluster1 = taxa[which(taxa_leadeigen$membership == 1)]
taxa_le_pca1 = prcomp(taxa_le_cluster1)
taxa_le_eigen1 = as.matrix(taxa_le_cluster1) %*% taxa_le_pca1$rotation
plot(taxa_le_pca1$sdev^2/sum(taxa_le_pca1$sdev^2), type="l", main = "Leading Eigen Cluster 1 PCs", 
     xlab="principle components", ylab="variance explained")

plot(taxa_le_eigen1[,1], taxa_le_eigen1[,2], main = "First cluster", 
     xlab="PCA1", ylab="PCA2")
summary(taxa_le_pca1)
taxa_le_cluster2 = taxa[which(taxa_leadeigen$membership == 2)]
taxa_le_pca2 = prcomp(taxa_le_cluster2)
taxa_le_eigen2 = as.matrix(taxa_le_cluster2) %*% taxa_le_pca2$rotation
plot(taxa_le_pca2$sdev^2/sum(taxa_le_pca2$sdev^2), type="l", main = "Leading Eigen Cluster 2 PCs", 
     xlab="principle components", ylab="variance explained")
plot(taxa_le_eigen2[,1], taxa_le_eigen2[,2], main = "Second cluster", 
     xlab="PCA1", ylab="PCA2")
summary(taxa_le_pca2)

taxa_le_cluster3= taxa[which(taxa_leadeigen$membership == 3)]
taxa_le_pca3 = prcomp(taxa_le_cluster3)
taxa_le_eigen3 = as.matrix(taxa_le_cluster3) %*% taxa_le_pca3$rotation
plot(taxa_le_pca3$sdev^2/sum(taxa_le_pca3$sdev^2), type="l", main = "Leading Eigen Cluster 3 PCs", 
     xlab="principle components", ylab="variance explained")
plot(taxa_le_eigen3[,1], taxa_le_eigen3[,2], main = "Third cluster", 
     xlab="PCA1", ylab="PCA2")
summary(taxa_le_pca3)

# taxa_le_cluster4= taxa[which(taxa_leadeigen$membership == 4)]
# taxa_le_pca4 = prcomp(taxa_le_cluster4)
# taxa_le_eigen4 = as.matrix(taxa_le_cluster4) %*% taxa_le_pca4$rotation
# plot(taxa_le_pca4$sdev^2/sum(taxa_le_pca4$sdev^2), type="l", main = "Leading Eigen Cluster 4 PCs", 
#      xlab="principle components", ylab="variance explained")
# 

# taxa_le_cluster5= taxa[which(taxa_leadeigen$membership == 5)]
# taxa_le_pca5 = prcomp(taxa_le_cluster5)
# taxa_le_eigen5 = as.matrix(taxa_le_cluster5) %*% taxa_le_pca5$rotation
# plot(taxa_le_pca5$sdev^2/sum(taxa_le_pca5$sdev^2), type="l", main = "Leading Eigen Cluster 5 PCs", 
#      xlab="principle components", ylab="variance explained")
```



# Linking Outcome

These are the pulmonary exacerbation events that have been entered into the database. As you will notice, there are often multiple events per person.

Not all patients or patient IDs on this list have a 16S microbiome sequence.

Note that pe_occurence=1 - is a definite yes; 2 is no (but was previously entered as having a PE due to some mild PE symptoms) and 3 is unknown. As a sensitivity analysis, you can exclude 2 and 3 to see if your deductions change. 

If pe_occurence is missing, you can exclude unless the comments indicate otherwise.

Dates are sometimes missing when the pe_date was unknown; use event_date instead,

```{r,echo=F}
taxa_short = taxa_avg %>% 
  filter(subject %in% unique(pulm$patient_id))
taxa_short$subject
pulm_new = pulm %>% group_by(patient_id) %>% 
  filter(patient_id %in% (taxa_avg$subject[unique(taxa_avg$subject) %in%
                                          unique(pulm$patient_id)]))
pulm_new$pe_occurance = ifelse(pulm_new$pe_occurance == 1, 1, 0)
pulm_new = pulm_new %>% 
  summarise(pulmonary = sum(pe_occurance, na.rm = T))
colnames(pulm_new) = c("subject", "pe")

taxa_merge = left_join(taxa_short, pulm_new, by = "subject")

View(taxa_merge[c("subject", "pe")])
```


# Poisson regression: first pcs for each of three clusters

```{r}
# Get eigentaxa
index_final = which(taxa_avg$subject %in% taxa_merge$subject)
## first pc for each cluster
taxa_pois = as.data.frame(cbind(taxa_le_eigen1[index_final,1], 
      taxa_le_eigen2[index_final,1],
      taxa_le_eigen3[index_final,1],
      taxa_merge$pe,
      ifelse(taxa_merge$pe == 0, 0, 1)))
colnames(taxa_pois) = c("taxa_eigen1", "taxa_eigen2", "taxa_eigen3",
                        "pe", "pe_bi")
summary(glm(pe ~ taxa_eigen1 + taxa_eigen2 + taxa_eigen3, family = "poisson", control = glm.control(maxit = 50),data = taxa_pois))
```

# Poisson regression: first pcs for each of three clusters

```{r}
# Get eigentaxa
index_final = which(taxa_avg$subject %in% taxa_merge$subject)
## first pc for each cluster
taxa_pois = as.data.frame(cbind(taxa_le_eigen1[index_final,1], taxa_le_eigen1[index_final,2], 
      taxa_le_eigen2[index_final,1],
      taxa_le_eigen2[index_final,2], 
      taxa_le_eigen3[index_final,1],
      taxa_le_eigen3[index_final,2], 
      taxa_merge$pe,
      ifelse(taxa_merge$pe == 0, 0, 1)))
colnames(taxa_pois) = c("taxa_eigen11","taxa_eigen12", "taxa_eigen21","taxa_eigen22", "taxa_eigen31","taxa_eigen32",
                        "pe", "pe_bi")
summary(glm(pe ~ taxa_eigen11 + taxa_eigen12+ taxa_eigen21 +  taxa_eigen22+ taxa_eigen31+ taxa_eigen32, family = "poisson", control = glm.control(maxit = 50),data = taxa_pois))
```

